---
title: "palmer_penguins_html"
author: "Charles Pugh"
date: "2025-03-27"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## **Palmer Penguins: Plots & Practice Analysis**
**Hypothesis:** Gentoo penguins have the largest flipper length of the penguin species witrhin the palmerpenguins dataset


### Setting up my environment
Notes: Setting up my R environment by loading the 'tidyverse' and 'palmerpenguins' packages:
```{r Loading packages}
library(tidyverse)
library(palmerpenguins)
```


### Previewing the data
```{r Using the head function to preview the data}
head(penguins)
```


```{r Using the colnames function to check the column names to see relevent data to confirm the hypothesis}
colnames(penguins)
```


### Plotting the data

The three charts below show a faceted comparison between flipper length and penguin species, separated and colour-coded by species type:

```{r The three charts below show a faceted comparison between flipper length and penguin species both separated and colour-coded by species type}
ggplot(data = penguins) +
  geom_bar(mapping = aes(x = flipper_length_mm, fill = species)) +
  facet_wrap(~species)
```

The chart below shows a comparison between flipper length and body mass color-coded for each species:

```{r The chart below shows a comparison between flipper length and body mass color-coded for each species}

ggplot(data = penguins) +

  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species))

```


The chart below shows a comparison between flipper length and species and is color-coded for each species:

```{r The chart below shows a faceted comparison between flipper length and penguin species, colour-coded by species type}
ggplot(data = penguins) +
  geom_point(mapping = aes(x = species, y = flipper_length_mm, color = species))
```


### Conclusion
It is clear that Gentoo species have the largest flippers, this confirms the hypothesis. They also have the largest body mass, suggesting a correlation between the two